Kindly Bent to Free Us

### 2019-11-11 kindly/soa.tex
SOA: Fixed typo.
SOA: Fixed typos.

Quill~\citep{DBLP:conf/icfp/Morris16} is an Haskell-like language with affine
types through the use of qualified types.
Quill does not exposes a kind language, but
uses typeclass-like mechanism and annotation on arrows.
It supports both a most general type for function composition and
principal type inference.
We were able to express the type signatures of all Quill examples
in \lang by leveraging kind inequalities.
Unlike Quill,
we provide a linearity-aware semantics and prove its soundness.
Quill does not support borrows.

@@

Linear and affine discipline have also been used in non-functional
settings, notably in the context of low-level imperative programming
and object systems. In particular, \lang uses borrows and regions
which were initially developed in the context of imperative and
object-oriented
programming~\citep{DBLP:conf/popl/BoylandR05,DBLP:conf/pldi/GrossmanMJHWC02}.

@@

\subsection{Type-system features}
%
\lang relies on constrained types
to introduce the kind inequalities required for linear types. In
particular, we use HM(X)~\citep{DBLP:journals/tapos/OderskySW99} which
allow to use constrained types in an ML-like language with complete
type inference.
HM(X) has been shown to be compatible with subtyping,
bounded universal and existentials~\citep{DBLP:conf/icfp/Simonet03},
GADTs~\citep{DBLP:journals/toplas/SimonetP07},
and there exists a syntactic soundness proof~\citep{DBLP:journals/entcs/SkalkaP02}.
These results make us confident that the system developed in \lang
could be applied to larger and more complex languages such as OCaml.

@@

The subtyping discipline used in \lang is very similar
to structural subtyping, where the only subtyping (or here, subkinding)
is at the leaves.
Such discipline is known to be friendly to inference and has be used in many
contexts, including OCaml itself, and has been combined
with constraints~\citep{DBLP:journals/tapos/OderskySW99,DBLP:conf/sas/TrifonovS96}.
In particular, Flowcaml~\citep{DBLP:conf/popl/PottierS02}
extends OCaml with security levels forming a lattice and supports type inference.
In \lang, we repurpose these ideas to check for linearity and affinity.
\citet{DBLP:conf/aplas/Simonet03} also presents various algorithmic
notions to solve lattice-based constraints, which we partially use
in our constraint solving algorithm. Our main contribution
is the design of a kind language specialized for checking linearity
and sufficiently simple to make
all simplification rules complete, which allows to keep type signatures simple.

>>>>

Quill~\citep{DBLP:conf/icfp/Morris16} is an Haskell-like language with affine
types through the use of qualified types.
Quill does not expose a kind language, but
uses typeclass-like mechanism and annotation on arrows.
It supports both a most general type for function composition and
principal type inference.
We were able to express the type signatures of all Quill examples
in \lang by leveraging kind inequalities.
Unlike Quill,
we provide a linearity-aware semantics and prove its soundness.
Quill does not support borrows.

@@

Linear and affine disciplines have also been used in non-functional
settings, notably in the context of low-level imperative programming
and object systems. In particular, \lang uses borrows and regions
which were initially developed in the context of imperative and
object-oriented
programming~\citep{DBLP:conf/popl/BoylandR05,DBLP:conf/pldi/GrossmanMJHWC02}.

@@

\subsection{Type-system features}
%
\lang relies on constrained types
to introduce the kind inequalities required for linear types. In
particular, we use HM(X)~\citep{DBLP:journals/tapos/OderskySW99} which
allows to use constrained types in an ML-like language with complete
type inference.
HM(X) has been shown to be compatible with subtyping,
bounded quantification and existentials~\citep{DBLP:conf/icfp/Simonet03},
GADTs~\citep{DBLP:journals/toplas/SimonetP07},
and there exists a syntactic soundness proof~\citep{DBLP:journals/entcs/SkalkaP02}.
These results make us confident that the system developed in \lang
could be applied to larger and more complex languages such as OCaml.

@@

The subtyping discipline used in \lang is very similar
to structural subtyping, where the only subtyping (or here, subkinding)
is at the leaves.
Such discipline is known to be friendly to inference and has been used in many
contexts, including OCaml itself, and has been combined
with constraints~\citep{DBLP:journals/tapos/OderskySW99,DBLP:conf/sas/TrifonovS96}.
In particular, Flowcaml~\citep{DBLP:conf/popl/PottierS02}
extends OCaml with security levels forming a lattice and supports type inference.
In \lang, we repurpose these ideas to check for linearity and affinity.
\citet{DBLP:conf/aplas/Simonet03} also presents various algorithmic
notions to solve lattice-based constraints, which we partially use
in our constraint solving algorithm. Our main contribution
is the design of a kind language specialized for checking linearity
and sufficiently simple to make
all simplification rules complete, which allows to keep type signatures simple.
